### 2020-10-30 ChapterBackground/chapterBackground.tex
Fixed: Proof-read corrections for first 3 pages of feedback response

\subsection{Consistency}
As online games are distributed (either between clients and a server, peers or multiple servers), inconsistent states can occur. Multiple peer/servers execute updates in parallel, however if these are executed out of order, such as due to messages being received out of order or messages lost entirely, inconsistent states can arise.  Different aspects of game state require different levels consistency and games often employ consistency resolution instead of inconsistency prevention \cite{P2PForMMOs}.

@@

Although there has been research utilising multiple servers to distribute the task of solving physics based problems (e.g., \cite{mashayekhi2018automatically}), there is no known literature describing real-time interactive physics exploiting the addition of servers to gain scalability. The closest work to this challenge is that carried out to seek scalability in terms of player numbers in online gaming in the field of Distributed Virtual Environments (DVEs), of which online gaming can be considered a subcategory.

Some key aspects to consider are the Goals and requirements of real-time DVEs, which are different from other distributed simulations. DVEs only need be "sufficiently realistic", which differs depending on context. This is known as selective fidelity, where different levels of realism are applied selectively to different aspects of the simulation, depending upon how much the aspect affects the outcome of the simulation. For example, in a ship simulator, realistic looking clouds will not contribute to simulator's effectiveness as a training device. DVEs must execute in real-time, unlike non-real-time simulations. Repeatability and synchronisation requirements are often relaxed as they may not contribute to the overall goal of the DVE, whereas they are essential for non-real-time simulations that require a high degree of accuracy. As a result of these different goals and requirements, many of the algorithms and techniques used in non-real-time simulations are not suitable for use in real-time DVEs \cite{fujimoto2000parallel}.

DVEs also share the same architecture categories as online gaming \cite{fujimoto2000parallel} i.e. centralized server (referred to as Client-Server in \ref{OnlineGaming}), Distributed server and distributed, serverless (referred to as peer-to-peer in \ref{OnlineGaming}). Online gaming also includes the hybrid peer-to-peer architecture.

A brief overview of some approaches to implementing systems for DVE will now be given.

\subsection{Distributed Interactive Simulation}
One example of an implementation of a DVE is Distributed Interactive Simulation (DIS), developed in the early '90s  \cite{fujimoto2000parallel}.
DIS followed the following design principles:
\begin{itemize}
	\item Autonomous simulation nodes.
	Each node is responsible for simulating one or more entities and generates messages, referred to as protocol data units (PDUs), to notify other nodes of updates to states of entities.
	\item Transmission of "ground truth" information. 
	Each node sends absolute truth concerning the state of the entities it represents.
	\item Transmission of state change information only. To economize on communications, simulation nodes only transmit changes in behaviour. (Referred to as delta-encoding and discussed in \ref{Delta-Encoding}).
	\item The use of "dead-reckoning", as discussed in \ref{ConsistencyRealTimePhysics}.
\end{itemize}

@@

\subsection{High Level Architecture}
HLA is a successor to DIS and was originally developed in the mid-'90s for military simulations. In HLA a distributed simulation comprises of individual simulators called federates and combined are referred to as a federation. The main goal behind HLA is to support interoperability and reuse of simulations (i.e. for use with heterogeneous simulations) \cite{fujimoto2000parallel}.

All entities in the simulation are represented in the HLA and each entity instance contains:
\begin{enumerate}
	\item A unique identifier
	\item Attributes that indicate those state variables and parameters of an object that are accessible to other objects
	\item Associations between objects.
\end{enumerate}
HLAs include a non-runtime and runtime component. The non-runtime component is responsible for specifying the definitions of the object types to be used by the federation. The runtime component is responsible for providing the federates with a means to interact with one another, known as the Run-time Infrastructure (RTI). Each federate is responsible for a set of entities and the state variables for entities are stored within the federates and not the runtime. Each federate is responsible for informing the RTI of any changes to state. Federates register interest with the RTI to receive updates for entities of which they are concerned \cite{fujimoto2000parallel}.

It should be noted that different attributes for a single entity can be owned by different federates and the RTI has no knowledge of the semantics of the information that it's transmitting. As a result of this lack of semantic knowledge, no optimisations (e.g. dead-reckoning) can be performed by the RTI and any techniques must be implemented only on the federates \cite{fujimoto2000parallel}.

HLA is ideal for heterogeneous distributed simulations and for re-use with different types of simulations. However, there are some disadvantages to this architecture:
\begin{enumerate}
	\item It increases latency between simulations as communication must go through the run time \item It does not allow for any optimisations to be performed in the run time as the run time is not aware of the semantics of the simulations.
\end{enumerate}

\subsection{SpatialOS}
SpatialOS uses a very similar architecture to HLA which also uses a runtime and 'workers' which are equivalent to federates \cite{SpatialOS}. A key difference between HLA and SpatialOS is that in SpatialOS the full entity data of all entities is stored within the run-time (as opposed to just attribute metadata and associations in HLA) and each worker stores a copy of a subset of the entities, which it is responsible for \cite{Introduc36:online}. Each worker communicates updates for the entities it is responsible for, which then get disseminated by the runtime to other workers with registered interest in those entities.

In the case of real-time physics, we can assume the use of either homogeneous simulations or at least simulations that will be simulating physics with a large degree of shared semantics between simulations (as all simulations are dealing with physical simulation). As a result of this, there is little to no advantage gained in using the HLA architecture or similar and direct communication between servers has the benefit of lower latency between servers.

>>>>

\subsection{Consistency}
As online games are distributed (either between clients and a server, peers or multiple servers), inconsistent states can occur. Multiple peers/servers execute updates in parallel, however if these are executed out of order, such as due to messages being received out of order or messages lost entirely, inconsistent states can arise.  Different aspects of game state require different levels of consistency, e.g. player life-death decisions require a high degree of consistency, whereas movements of other players have little effect on gameplay (in games like World of Warcraft) and so only require a low degree of consistency \cite{Savery2014Consistency}. In addition, games often employ consistency resolution instead of inconsistency prevention/consistency maintenance \cite{P2PForMMOs}.

Four categories of algorithms for consistency maintenance techniques for networked games are identified in \cite{Savery2014Consistency}. The categories are as follows:
\begin{itemize}
	\item Delay the processing of local events. Examples include Bucket Synchronisation \cite{BucketSynchronisation1998} and Local Lag \cite{LocalLag2004}, in which local actions are grouped into buckets of time or delayed to allow for simultaneous execution on all clients.
	\item Alter the time frame for remote entities e.g. Remote Lag \cite{Bernier2003LatencyCM}, in which local actions are applied immediately, but remote actions are delayed, allowing for interpolation (instead of relying on prediction), and Local Perception Filters \cite{PerceptionFilter}, a refinement on Remote Lag, in which, instead of applying a fixed delay to remotely controlled entities (as in Remote Lag), a variable delay is applied determined by the position of the remote entity relative to the player.
	\item Absolute consistency e.g. Locking and Serialisation \cite{coulouris2005distributed}, similar to techniques used in transactional databases, and TimeWarp \cite{LocalLag2004}, in which each client keeps a list of commands and states ordered by time and when a command arrives late or out of order, the command is inserted into the list with the correct ordering and the commands are then applied in order, starting from the state immediately preceding the new command, until the state reaches the current time.
	\item Predictive algorithms, such as dead-reckoning (discussed further in section \ref{ConsistencyRealTimePhysics}), in which future positions of remote entities are predicted based on prior data, such as position and velocity.
\end{itemize}

Consistency resolution can divided into the following two categories:
\begin{itemize}
\item Decision-making. Determining the outcome of decisions in the presence of inconsistency. The decisions can be determined by choosing a perspective, i.e. the server, in which the server maintains the whole state necessary to determine the decision; source client, the perspective of the client that initiates the action (e.g. shoots an enemy player); and target client, the perspective of the client that the action targets (e.g. target of an enemy player shooting) \cite{Savery2014Consistency}.

\item Error Repair. Repairing inconsistencies as they are discovered. The three basic approaches to error repair are: 
	Correct Immediately, in which remote entities are immediately updated to the latest received positions; Smoothly Correct, where remote entities' positions are interpolated over time to the latest position; and Tolerate, where if the inconsistency has no effect on gameplay, the inconsistency is not corrected \cite{Savery2014Consistency}.
\end{itemize}

@@

Although there has been research utilising multiple servers to distribute the task of solving physics-based problems (e.g., \cite{mashayekhi2018automatically}), there is no known literature describing real-time interactive physics exploiting the addition of servers to gain scalability. The closest work to this challenge is that carried out to seek scalability in terms of player numbers in online gaming in the field of Distributed Virtual Environments (DVEs), of which online gaming can be considered a subcategory.

Some key aspects to consider are the goals and requirements of real-time DVEs, which are different from other distributed simulations. DVEs only need to be "sufficiently realistic", which differs depending on context. This is known as selective fidelity, where different levels of realism are applied selectively to different aspects of the simulation, depending upon how much the aspect affects the outcome of the simulation. For example, in a ship simulator, realistic-looking clouds will not contribute to the simulator's effectiveness as a training device. DVEs must execute in real-time, unlike non-real-time simulations. Repeatability and synchronisation requirements are often relaxed as they may not contribute to the overall goal of the DVE, whereas they are essential for non-real-time simulations that require a high degree of accuracy. As a result of these different goals and requirements, many of the algorithms and techniques used in non-real-time simulations are not suitable for use in real-time DVEs \cite{fujimoto2000parallel}.

DVEs also share the same architecture categories as online gaming \cite{fujimoto2000parallel} i.e. centralised server (referred to as client-server in \ref{OnlineGaming}), distributed server and distributed serverless (referred to as peer-to-peer in \ref{OnlineGaming}). Online gaming also includes the hybrid peer-to-peer architecture.

A brief overview of some approaches to implementing systems for DVEs will now be given.

\subsection{Distributed Interactive Simulation}
One example of an implementation of a DVE is Distributed Interactive Simulation (DIS), developed in the early '90s  \cite{fujimoto2000parallel}.
DIS followed the following design principles:
\begin{itemize}
	\item Autonomous simulation nodes.
	Each node is responsible for simulating one or more entities and generates messages, referred to as Protocol Data Units (PDUs), to notify other nodes of updates to states of entities.
	\item Transmission of "ground truth" information. 
	Each node sends absolute truth concerning the state of the entities it represents.
	\item Transmission of state change information only. To economise on communications, simulation nodes only transmit changes in behaviour. (Referred to as delta-encoding and discussed in \ref{Delta-Encoding}).
	\item The use of "dead-reckoning", as discussed in \ref{ConsistencyRealTimePhysics}.
\end{itemize}

@@

\subsection{High Level Architecture}
HLA is a successor to DIS and was originally developed in the mid-'90s for military simulations. In HLA a distributed simulation comprises of individual simulators called federates, which combined are referred to as a federation. The main goal behind HLA is to support interoperability and reuse of simulations (i.e. for use with heterogeneous simulations) \cite{fujimoto2000parallel}.

All entities in the simulation are represented in the HLA and each entity instance contains:
\begin{enumerate}
	\item A unique identifier
	\item Attributes that indicate those state variables and parameters of an object that are accessible to other objects
	\item Associations between objects.
\end{enumerate}
HLAs include a non-runtime and runtime component. The non-runtime component is responsible for specifying the definitions of the object types to be used by the federation. The runtime component is responsible for providing the federates with a means to interact with one another, known as the Run-time Infrastructure (RTI). Each federate is responsible for a set of entities and the state variables for entities are stored within the federates and not the RTI. Each federate is responsible for informing the RTI of any changes to state. Federates register interest with the RTI to receive updates for entities for which they are concerned \cite{fujimoto2000parallel}.

It should be noted that different attributes for a single entity can be owned by different federates and the RTI has no knowledge of the semantics of the information that it is transmitting. As a result of this lack of semantic knowledge, no optimisations (e.g. dead-reckoning) can be performed by the RTI and any techniques must be implemented only on the federates \cite{fujimoto2000parallel}.

HLA is ideal for heterogeneous distributed simulations and for re-use with different types of simulations. However, there are some disadvantages to this architecture:
\begin{enumerate}
	\item It increases latency between simulations as communication must go through the RTI.
	\item It does not allow for any optimisations to be performed in the RTI as the RTI is not aware of the semantics of the simulations.
\end{enumerate}

\subsection{SpatialOS}
SpatialOS uses a very similar architecture to HLA which also uses an RTI and 'workers' which are equivalent to federates \cite{SpatialOS}. A key difference between HLA and SpatialOS is that in SpatialOS the full entity data of all entities is stored within the RTI (as opposed to just attribute metadata and associations in HLA) and each worker stores a copy of a subset of the entities, which it is responsible for \cite{Introduc36:online}. Each worker communicates updates for the entities it is responsible for, which then get disseminated by the RTI to other workers with registered interest in those entities.

In the case of real-time physics, we can assume the use of either homogeneous simulations or at least simulations that will be simulating physics with a large degree of shared semantics between simulations (as all simulations are dealing with physical simulation). As a result of this, there is little to no advantage gained in using HLA or similar architectures, furthermore direct communication between servers has the benefit of lower latency between servers.
